Secure Composition \\ of Robust and Optimising Compilers

### 2023-09-27 csf/main.tex
crunch TODOs

  \item %
        %\Cref{sec:compcomp} studies different forms of secure compiler compositions.
        This paper takes the secure compilation framework of Abate et al.~\cite{abate2019jour} and extends it to reason about the security of all different known forms of compiler composition (\Cref{sec:compcomp}).
        %Concretely, it investigates upper and lower compositions, as well as the sequential composition.
        For this, this paper studies sequential compiler composition as well as compilers with multiple input languages or multiple output ones, as used in existing compilation chains.
        %
        This paper proves that starting from two compilers that preserve two (possibly distinct) properties, their composition preserves the intersection of those properties.
        % For example, a compiler that is secure with respect to \gls*{tms} can be composed with a compiler that is secure with respect to \gls*{sms}, resulting in a compiler that is secure with respect to \gls*{ms}.
        % 
        Finally, this paper proves that the order of composition of sequential compiler passes is irrelevant for the resulting security.
        % 
        This is crucial for reordering optimisation passes and thus generating secure and efficient code.
      	\MP{
      		doublecheck.
      		also, the corollary .. ?
      	}

@@

\MP{

@@

@@

\subsection{Shared Language Definitions}\label{subsec:cs:defs}
\MPin{
	note to fix as agreed upon
	+ red syntax
	+ green syntax
}
\begin{gather*}
  \begin{aligned}
    \mi{(Expressions)}~&\expr\bnfdef\ x \mid \valueexpr \mid \lbinop{\expr[_{1}]}{\expr[_{2}]} \mid \lget{x}{\expr}\\
                       &\mid \llet{x}{\lnew{\expr[_{1}]}{\expr[_{2}]}}{\expr[_{3}]} \mid \ldelete{x}\\
                       &\mid \lset{x}{\expr[_{1}]}{\expr[_{2}]} \mid \lpair{\expr[_{1}]}{\expr[_{2}]} \mid \lproja{\expr} \mid \lprojb{\expr}\\
                       &\mid \llet{x}{\expr[_{1}]}{\expr[_{2}]} \mid \lreturn{\expr} \mid \lcall{g}{\expr}\\
                       &\mid \lifz{\expr}{\expr[_{1}]}{\expr[_{2}]} \mid \labort
  \end{aligned}
  \\
  \begin{aligned}
  \mi{(Functions)}~\function\bnfdef&\ \lfunction{\textit{foo}}{x}{e}
  \end{aligned}
  \\
  \begin{aligned}
  \mi{(Libraries)}~\library\bnfdef&\ \hole{\cdot} \mid \function,\library\hspace{0.5cm}
  \mi{(Programs)}~\prog{\library_{\ctx}}{\library_{\comp}}
  \end{aligned}
\end{gather*}

@@

\begin{theorem}[$\src{L_{\tmssafe}}$-programs are \gls*{tms}]\label{thm:wt:tms}
  For any $\src{\library_{\comp}}$, $\rsat{\src{\library_{\comp}}}{\tmssafe}$ \Coqed
\end{theorem}

@@

\vspace{-2em}
\begin{center}
  \begin{align*}
    \mi{(Expressions)}~\trg{\expr}\bnfdef& \cdots \mid \trg{\lispoisoned{x}} \mid \trg{\lhast{\expr}{\type}} \\
  \end{align*}
\end{center}
\vspace{-2em}

$\Ltrg$ extends the syntax presented earlier (\Cref{subsec:cs:defs}) with dynamic typechecks $\trg{\lhast{\expr}{\type}}$ and a way to inspect poison tags $\trg{\lispoisoned{x}}$ in the metadata of pointers.
For valid pointers ($\trg{\poisonless}$) bound to variable $\trg{x}$, the check $\trg{\lispoisoned{x}}$ yields $\trg{1}$.
If the array bound to $\trg{x}$ was deallocated, i.e., has been poisoned ($\trg{\poisoned}$), the check $\trg{\lispoisoned{x}}$ evaluates to $\trg{0}$.

@@

\vspace{-1em}
\begin{gather*}
  \begin{aligned}
    \mi{(Expressions)}~\obj{\expr}\bnfdef&\ \obj{n^{\securitytag}} \mid \cdots \mid \obj{\llet{x^{\securitytag}}{\expr[_{1}]}{\expr[_{2}]}} \mid \cdots\\&\mid \obj{\lwrdoit{D}} \mid \obj{\lrddoit{x}{\expr}}
  \end{aligned} 
  \\
  \begin{aligned}
    \mi{(States)}~\obj{\statevar}&\bnfdef \obj{\tup{\cfstate}{\sandboxtag,D,\memstate}}
     &
    \mi{(DIT\ Flag)}~\obj{D}&\bnfdef \obj{ON} \mid \obj{OFF}
  \end{aligned}
\end{gather*}

$\Lscct$ extends $\Lms$ (\Cref{subsec:lms}) with a way to write to a model specific register that controls a data (operand) independent timing mode, a feature that is present in both Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman} processors.
To this end, states are extended with the value of the register, which is initially set to be not active.
If the register is marked active, the intuition is that no secrets can appear on specification traces.
If the register is marked inactive, secrets may appear on traces.
For the other languages seen earlier, the mode is intuitively always-on, i.e., the mode of execution always uses data independent timing.
The language also adds user-annotations to values and variables to know their secrecy $\obj{\securitytag}$, which is either high $\obj{\lock}$ or low $\obj{\unlock}$.
Security tags $\obj{\securitytag}$ are on the usual secrecy lattice, where $\obj{\lock}\le\obj{\securitytag}$ and $\obj{\securitytag}\le\obj{\unlock}$.

\vspace{-1em}
% \begin{center}
{\small
  \begin{align*}
    \mi{(Pre\text{-}Events)}~\obj{\event[_{b}]}\bnfdef&\ \cdots \mid \obj{\widehat{Get}\ \loc\ \valueexpr} \mid \obj{\widehat{Set}\ \loc\ \valueexpr\ \valueexpr[']}
    % \\
                                                      % &
                                                      \mid \obj{Branch\ n} \mid \obj{Binop\ n}
  \end{align*}
% \end{center}
}

To prevent secrets from leaking but still enable reasoning about memory safety, $\Lscct$ extends pre-events with $\obj{\widehat{Get}\ \loc\ \valueexpr}$ and $\obj{\widehat{Set}\ \loc\ \valueexpr}$.
These indicate reads from and writes to memory without leaking secret information involved in the access.
Moreover, the language extends pre-events with $\obj{\ev{Branch}\ n}$ and $\obj{\ev{Binop}\ n}$ that are emitted when evaluating a branch or certain binary expressions, such as division, respectively, whenever the data independent timing mode is inactive.
The following rules demonstrate how this is handled semantically.

>>>>

  \item %
        %\Cref{sec:compcomp} studies different forms of secure compiler compositions.
        This paper takes the secure compilation framework of Abate et al.~\cite{abate2019jour} and extends it to reason about the security of all different known forms of compiler composition (\Cref{sec:compcomp}).
        %Concretely, it investigates upper and lower compositions, as well as the sequential composition.
        For this, this paper studies sequential compiler composition as well as compilers with multiple input languages or multiple output ones, as used in existing compilation chains.
        %
        This paper proves that starting from two compilers that preserve two (possibly distinct) properties, their composition preserves the intersection of those properties.
        % For example, a compiler that is secure with respect to \gls*{tms} can be composed with a compiler that is secure with respect to \gls*{sms}, resulting in a compiler that is secure with respect to \gls*{ms}.
        % 
        From this, the paper proves that the order of composition of sequential compiler passes is irrelevant for the resulting security.
        % 
        This is crucial for reordering individual optimisation passes and thus generating secure and efficient code.
        %
        Finally, this paper extends parts of the compositionality framework to support different trace-models between compilation passes.

@@

\MPin{

@@

\MKin{ Optimisations that remove elements from traces are not necessarily sound,
 e.g., remove all deallocations to "optimise" memory management by offloading
 the work to the OS.

@@

\subsection{Shared Language Definitions}\label{subsec:cs:defs}

All presented programming languages share a common fragment. 
Programs ($\prog{\library_{\ctx}}{\library_{\comp}}$) consist of libraries ($\library$), which are lists of functions ($\function$).
Functions ($\lfunction{\textit{foo}}{x}{\expr}$) have a single argument $x$ and an expression ($\expr$) as a body which may have $x$ as a free variable.
Expressions can be variables ($x$), values ($\valueexpr$), binary operations ($\lbinop{\expr[_1]}{\expr[_2]}$), reading from ($\lget{x}{\expr}$) and writing to ($\lset{x}{\expr[_1]}{\expr[_2]}$) memory, memory allocation ($\llet{x}{\lnew{\expr[_1]}{\expr[_2]}}{\expr[_3]}$) as well as deallocation ($\ldelete{x}$), let expressions ($\llet{x}{\expr[_1]}{\expr[_2]}$), pairs ($\lpair{\expr[_1]}{\expr[_2]}$), their first ($\lproja{\expr}$) and second projection ($\lprojb{\expr}$), conditionals ($\lifz{\expr[_1]}{\expr[_2]}{\expr[_3]}$), function calls ($\lcall{\texttt{foo}}{\expr}$), returns ($\lreturn{\expr}$), and an aborting expression ($\labort$).

@@

\begin{theorem}[$\src{L_{\tmssafe}}$-programs are \gls*{tms}]\label{thm:wt:tms}
  $\;$
  \begin{nscenter}
  For any $\src{\library_{\comp}}$, $\rsat{\src{\library_{\comp}}}{\tmssafe}$ \Coqed
  \end{nscenter}
\end{theorem}

@@

$\Ltrg$ extends the syntax presented earlier (\Cref{subsec:cs:defs}) with dynamic typechecks ($\trg{\lhast{\expr}{\type}}$) and a way to inspect poison tags ($\trg{\lispoisoned{x}}$) in the metadata of pointers.
For valid pointers ($\trg{\poisonless}$) bound to variable $\trg{x}$, the check $\trg{\lispoisoned{x}}$ yields $\trg{1}$.
If the array bound to $\trg{x}$ was deallocated, i.e., has been poisoned ($\trg{\poisoned}$), the check $\trg{\lispoisoned{x}}$ evaluates to $\trg{0}$.

@@

$\Lscct$ extends $\Lms$ (\Cref{subsec:lms}) with a way to read from ($\obj{\lrddoit{x}{\expr}}$) and to write to ($\obj{\lwrdoit{D}}$) a model specific register that controls a data (operand) independent timing mode, a feature that is present in both Arm~\cite[p.~543]{arm-refman} and Intel~\cite[p.~80]{intel-refman} processors.
To this end, states are extended with the value of the register ($\obj{D}$), which is initially set to be not active ($\obj{OFF}$).
If the register is marked active ($\obj{ON}$), the intuition is that no secrets can appear on specification traces.
If the register is marked inactive, secrets may appear on traces.
For the other languages seen earlier, the mode is intuitively always-on, i.e., the mode of execution always uses data independent timing.
The language also adds user-annotations to values and variables to know their secrecy $\obj{\securitytag}$, which is either high $\obj{\lock}$ or low $\obj{\unlock}$.
Security tags $\obj{\securitytag}$ are on the usual secrecy lattice, where $\obj{\lock}\le\obj{\securitytag}$ and $\obj{\securitytag}\le\obj{\unlock}$.
To prevent secrets from leaking but still enable reasoning about memory safety, $\Lscct$ extends pre-events with $\obj{\widehat{Get}\ \loc\ \valueexpr}$ and $\obj{\widehat{Set}\ \loc\ \valueexpr}$.
These indicate reads from and writes to memory without leaking secret information involved in the access.
Moreover, the language extends pre-events with $\obj{\ev{Branch}\ n}$ and $\obj{\ev{Binop}\ n}$ that are emitted when evaluating a branch or certain binary expressions, such as division, respectively, whenever the data independent timing mode is inactive.
The following rules demonstrate how this is handled semantically.
